On the uniqueness and stability of dictionaries \\ for sparse representation of noisy signals

### 2019-05-08 manuscripts/IEEE/final_submission.tex
final through page 3

%\cite{li2004analysis, Georgiev05, Aharon06, Hillar15}
We should stress the point here that framing the problem in terms of hypergraphs has allowed us to show, unlike in previous works on the subject, that the matrix $\mathbf{A}$ need not necessarily satisfy \eqref{SparkCondition} to be recoverable from data. As an example, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ with $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$ and take $\mathcal{H}$ to be all consecutive pairs of indices $1, \ldots ,6$ arranged in cyclic order. Then for $k=2$, the matrix $\mathbf{A}$ fails to satisfy \eqref{SparkCondition} while still satisfying the first set of assumptions in Thm.~\ref{DeterministicUniquenessTheorem}.%, hence guaranteeing \eqref{Cstable} . %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
%This weakening allows for a practical (polynomial) amount of data to still guarantee a stable dictionary in the case where the set of sparse supports for the $\mathbf{\overline x}_i$ is known to have a size that grows polynomially in $m$ and $k$.

@@

We claim that a similar phenomenon applies to datasets of vectors with a stable sparse representation. Briefly, following the same procedure as in \cite[Sec.~IV]{Hillar15}, for $k < m$ and $n \geq \min(2k, m)$ we may consider the ``symbolic'' dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by an indeterminate $n \times m$ matrix $\mathbf{A}$ and $m$-dimensional $k$-sparse vectors $\mathbf{x}_1, \ldots, \mathbf{x}_N$ indeterminate within their supports, which form a regular hypergraph $\mathcal{H} \subseteq {[m] \choose k}$ satisfying the SIP. Restricting \mbox{$(k-1){m \choose k} + 1$} indeterminate $\mathbf{x}_i$ to each support in $\mathcal{H}$, and letting $\textbf{M}$ be the $n \times N$ matrix with columns $\mathbf{A}\mathbf{x}_i$, it can be checked that when $f(\mathbf{M}) \neq 0$ for a substitution of real numbers for the indeterminates, all of the assumptions on $\mathbf{A}$ and $\mathbf{x}_i$ in Thm.~\ref{DeterministicUniquenessTheorem} are satisfied; in particular, $\mathbf{A}$ satisfies \eqref{SparkCondition}. We therefore have the following result.

@@

To extend this observation to arbitrary probability distributions, note that if a set of measure spaces $\{(X_{\ell}, \Sigma_{\ell}, \nu_{\ell})\}_{\ell=1}^p$ has $\nu_{\ell}$ absolutely continuous with respect to $\mu$ for all $\ell \in [p]$, where $\mu$ is the standard Borel measure on $\mathbb{R}$, then the product measure $\prod_{\ell=1}^p \nu_{\ell}$ is absolutely continuous with respect to the standard Borel product measure on $\mathbb{R}^p$ (e.g., see \cite{folland2013real}).  This fact combined with Thm.~\ref{robustPolythm} implies the following.

>>>>

%\cite{li2004analysis, Georgiev05, Aharon06, Hillar15}
We should stress the point here that framing the problem in terms of hypergraphs has allowed us to show, unlike in previous works on the subject, that the matrix $\mathbf{A}$ need not necessarily satisfy \eqref{SparkCondition} to be recoverable from data. As an example, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ with $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$ and take $\mathcal{H}$ to be all consecutive pairs of indices $1, \ldots ,6$ arranged in cyclic order. Then for $k=2$, the matrix $\mathbf{A}$ fails to satisfy \eqref{SparkCondition} while still obeying the first set of assumptions in Thm.~\ref{DeterministicUniquenessTheorem}.%, hence guaranteeing \eqref{Cstable} . %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
%This weakening allows for a practical (polynomial) amount of data to still guarantee a stable dictionary in the case where the set of sparse supports for the $\mathbf{\overline x}_i$ is known to have a size that grows polynomially in $m$ and $k$.

@@

We claim that a similar phenomenon applies to datasets of vectors with a stable sparse representation. Briefly, following the same procedure as in \cite[Sec.~IV]{Hillar15}, for $k < m$ and $n \geq \min(2k, m)$ we may consider the ``symbolic'' dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by an indeterminate $n \times m$ matrix $\mathbf{A}$ and $m$-dimensional $k$-sparse vectors $\mathbf{x}_1, \ldots, \mathbf{x}_N$ indeterminate within their supports, which form a regular hypergraph $\mathcal{H} \subseteq {[m] \choose k}$ satisfying the SIP. Restricting \mbox{$(k-1){m \choose k} + 1$} indeterminate $\mathbf{x}_i$ to each support in $\mathcal{H}$, and letting $\textbf{M}$ be the $n \times N$ matrix with columns $\mathbf{A}\mathbf{x}_i$, it can be checked that when $f(\mathbf{M}) \neq 0$ for a substitution of real numbers for the indeterminates, all of the assumptions on $\mathbf{A}$ and $\mathbf{x}_i$ in Thm.~\ref{DeterministicUniquenessTheorem} are satisfied. We therefore have the following result.  % ; in particular, $\mathbf{A}$ satisfies \eqref{SparkCondition}

@@

To extend this observation to arbitrary probability distributions, note that if a set of measure spaces $\{(X_{\ell}, \nu_{\ell})\}_{\ell=1}^p$ has $\nu_{\ell}$ absolutely continuous with respect to $\mu$ for all $\ell \in [p]$, where $\mu$ is the standard Borel measure on $\mathbb{R}$, then the product measure $\prod_{\ell=1}^p \nu_{\ell}$ is also absolutely continuous with respect to the standard Borel product measure on $\mathbb{R}^p$ (e.g., see \cite{folland2013real}).  This fact combined with Thm.~\ref{robustPolythm} implies the following.
